The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-02-02 paper/paper.tex
another method

Furthermore, metacal provides a solution to the pernicious and general problem of
\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
that are
highly covariant with those on other galaxy properties, including size and, most importantly, flux/
magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
depending on their shear.  Within metacal, we can solve this problem by performing the cut on both
sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
variants.  In DES the corrected biases were found up to around 3\%, far larger than
the requirements for the current generation of surveys \cite{des-y1-cat}.

@@

Random forests \cite{breiman2001} train a collection of individual \emph{decision trees},
each of which consists of a set of bifurcating comparisons in which different features in
the data (in this case, magnitudes and errors) are compared to criterion values to choose
which branch to follow.  Each ``leaf'' (end comparison) selects a classification bin for
an input galaxy, and the tree is trained to choose comparison features and comparison values
which maximally split the discrimination at each step, and in final leaf purity.

@@

\subsection{ {\sc ComplexSOM} }

@@

GPzBin is a simple extension to GPz for the problem of tomographic binning. For
a given number of tomographic bins it selects redshift bin edges $z_{i}$ such
that there are an equal number of training galaxies in each bin. Testing data
galaxies then are assigned to the bin corresponding to their GPz predicted
photometric redshift. The code allows for two selection criteria for removing
galaxies for which it was not possible to make a good photometric redshift
prediction: 1) cutting galaxies close to the boundaries of bins based on an
`edge strictness' parameter $U$ which removes galaxies where $\mu\pm U \times
\sigma \lessgtr z_i$ and 2) cutting galaxies based on the degree $E$ to which
GPz is having to extrapolate to make the redshift prediction, removing galaxies
with $\nu>E\sigma^2$ (see Figure 12 of \cite{Hatfield2020}). Cutting galaxies
removes some signal at the possible benefit of improving binning purity. As the
training and test data were sampled from the same distribution in this data
challenge we might not expect removing galaxies to give huge improvements in
binning quality, but cuts based on $E$ and $U$ might become more useful in
future studies where the training and test data have substantially different
colour-magnitude distributions.

When used in GPzBin for this data challenge the GPz settings
in Table \ref{table-settings} were used (see
\cite{Almosallam2016a,Almosallam2016b} for precise definitions and
interpretations).

@@

In order to determine optimal centroid positions, clustering is framed as a
gradient descent problem that seeks to maximize the requested metric: either the
SNR or the FoM. This approach of identifying clusters using gradient descent is
inspired by the similar approach to K-means clustering in
\cite{NIPS1994_a1140a3d}.

@@

\begin{itemize}
    \item Splitting classification from nominal bin choice
    \item Infrastructure for checking submissions
    \item Guidance for users regarding caching and batch processing
    \item Timing
    \item Training set size
    \item Advertisement
\end{itemize}

>>>>

Furthermore, metacal provides a solution to the pernicious and general problem of
\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
that are
highly covariant with those on other galaxy properties, including size and, most importantly, flux/
magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
depending on their shear.  Within metacal, we can solve this problem by performing the cut on both
sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
variants.  In DES the corrected biases were found up to around 3\%, far larger than
the requirements for the current generation of surveys \citep{des-y1-cat}.

@@

Random forests \citep{breiman2001} train a collection of individual \emph{decision trees},
each of which consists of a set of bifurcating comparisons in which different features in
the data (in this case, magnitudes and errors) are compared to criterion values to choose
which branch to follow.  Each ``leaf'' (end comparison) selects a classification bin for
an input galaxy, and the tree is trained to choose comparison features and comparison values
which maximally split the discrimination at each step, and in final leaf purity.

@@

\subsection{ {\sc ComplexSOM} }
  The {\tt ComplexSOM} method implements an additional optimization layer on the methodology used by {\tt SimpleSOM}.

  Let ${\sf C}^{\rm group}_\ell$ be the matrix containing all auto- and cross-power spectra between members of a large set of galaxy samples (called ``groups'' here). We can compress this large set into a smaller set of samples (called ``bins'' here), by combining different groups. Let ${\sf A}$ be the assignment matrix determining these bins, such that $A_{\alpha,i}=1$ if group $i$ is assigned to bin $\alpha$, and 0 otherwise. In that case, the redshift distribution of bin $\alpha$ is given by $N_\alpha(z)=\sum_i A_{\alpha,i}n_i(z)$, where $n_i(z)$ is the redshift distribution of group $i$. Defining the normalized assignment matrix ${\sf B}$ as
  \begin{equation}
    B_{\alpha,i}=A_{\alpha,i}\frac{\int dz\,n_i(z)}{\sum_j A_{\alpha,j}\int dz\,n_j(z)},
  \end{equation}
  ${\sf C}^{\rm group}_\ell$ is related to the matrix of cross-power spectra between pairs of bins via:
  \begin{equation}
    {\sf C}^{\rm bin}_\ell={\sf B}\,{\sf C}^{\rm group}_\ell\,{\sf B}^T.
  \end{equation}

  The rationale behind the {\tt ComplexSOM} method is based on the observation that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are very fast. Thus, given an initial set of groups characterizing the distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$ can be precomputed once, and used to find the optimal assignment matrix ${\sf B}$ that maximizes the figure of merit in Eq. {\bf reference to FoM equation}, which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as
  \begin{equation}\label{eq:fom_tr}
    F_{\theta\theta'}=\sum_\ell f_{\rm sky}\frac{2\ell+1}{2}{\rm Tr}\left[\partial_\theta{\sf C}^{\rm bin}_\ell({\sf C}^{\rm bin}_\ell)^{-1}\partial_{\theta'}{\sf C}^{\rm bin}_\ell({\sf C}_\ell^{\rm bin})^{-1}\right].
  \end{equation}

  Note that the general problem of finding the matrix ${\sf B}$ that optimally compresses the information on a given parameter has an analytical solution in terms of Karnhunen-L\^oeve eigenvectors \citep{astro-ph/9603021}. However, the assignment matrix used in this case has the additional constraints that it must be positive, discrete and binary (i.e. groups are either in a bin or they are not). Thus, finding the absolute maximum figure of merit would involve a brute-force evaluation of all matrix elements $A_{\alpha,j}$, which is an intractable $N_{\rm bin}^{N_{\rm group}}$ problem. We thus parametrize ${\sf A}$ in terms of a small set of continuous parameters, for which standard minimization routines can be used. In particular we use $N_{\rm bin}-1$ parameters, given by the edges of the redshift bins, and a group is assigned to a bin if its mean redshift falls within its edges. We also explored the possibility of ``trashing'' groups if less than a fraction $p_{\rm in}$ of its redshift distribution fell within its preassigned bin, treating $p_{\rm in}$ as an additional free parameter. We found that adding this extra freedom did not translate into a significant improvement in the final level of data compression.

  Once the SOM has been generated, the {\tt ComplexSOM} method proceeds as follows. First, the full SOM is distilled into a smaller set of ${\cal O}(100)$ groups based by combining SOM cells with similar moments of their redshift distributions, thus ensuring that this process does not induce any catastrophic $N(z)$ broadening {\bf you may wanna refer to the {\tt SimpleSOM} method here}. The large ${\sf C}^{\rm group}_\ell$ is then precomputed for the initial set of groups. Finally, we find the optimal assignment matrix, defined in terms of the free redshift-edge parameters, by maximizing the figure of merit in Eq. \ref{eq:fom_tr} using Powell's minimization method \citep{10.1093/comjnl/7.2.155}.

@@

GPzBin is a simple extension to GPz for the problem of tomographic binning. For
a given number of tomographic bins it selects redshift bin edges $z_{i}$ such
that there are an equal number of training galaxies in each bin. Testing data
galaxies then are assigned to the bin corresponding to their GPz predicted
photometric redshift. The code allows for two selection criteria for removing
galaxies for which it was not possible to make a good photometric redshift
prediction: 1) cutting galaxies close to the boundaries of bins based on an
`edge strictness' parameter $U$ which removes galaxies where $\mu\pm U \times
\sigma \lessgtr z_i$ and 2) cutting galaxies based on the degree $E$ to which
GPz is having to extrapolate to make the redshift prediction, removing galaxies
with $\nu>E\sigma^2$ (see Figure 12 of \citealp{Hatfield2020}). Cutting galaxies
removes some signal at the possible benefit of improving binning purity. As the
training and test data were sampled from the same distribution in this data
challenge we might not expect removing galaxies to give huge improvements in
binning quality, but cuts based on $E$ and $U$ might become more useful in
future studies where the training and test data have substantially different
colour-magnitude distributions.

When used in GPzBin for this data challenge the GPz settings
in Table \ref{table-settings} were used (see
\citealp{Almosallam2016a,Almosallam2016b} for precise definitions and
interpretations).

@@

In order to determine optimal centroid positions, clustering is framed as a
gradient descent problem that seeks to maximize the requested metric: either the
SNR or the FoM. This approach of identifying clusters using gradient descent is
inspired by the similar approach to K-means clustering in
\citet{NIPS1994_a1140a3d}.

@@

\subsection{Splitting nominal bin choice from assignment}
There were two ways to obtain good scores in the challenge: either one could choose the best
nominal bin edges, or find the best method to put galaxies into those bins once they are chosen.
These two are not completely disconnected: some theoretically powerful choices of nominal edges
could prove impossible to assign in practice.  But it would still have been useful to separate
the challenge into two separate parts, fixing the edges while optimizing classifiers, and vice
versa.   We can retrospectively determine this since we understand the classifiers, but the picture
would have been clearer doing so in advance.

\subsection{Infrastructure}

\begin{itemize}
    \item 
    \item Infrastructure for checking submissions
    \item Guidance for users regarding caching and batch processing
    \item Timing
    \item Training set size
    \item Advertisement
\end{itemize}
